---
title: "terrorism_poverty"
output: html_notebook
---

# Data Load
## Import Necessary Libraries
Load the R libraries required for data manipulation, visualization, and querying. Libraries like dplyr and tidyr are used for data cleaning and transformation, ggplot2 for plotting, sqldf for SQL queries, and rnaturalearth for map data.
```{r}
# Load the necessary libraries
library(DBI)
library(RSQLite)
library(dbplyr)
library(dplyr)
library(sqldf)
library(tidyr)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
```

## Load the Data
Here, you read the datasets for terrorism and poverty from CSV files located on your local drive into R. read.csv() loads these files into data frames named terrorism_data and poverty_data.
```{r}
terrorism_data <- read.csv("D:/globalterrorism.csv")
poverty_data <- read.csv("D:/pip_dataset.csv")
```

# Data Exploratory
## Identification
This chunk identifies and lists the unique countries present in each dataset. It helps to understand which countries are represented in both datasets and which are exclusive to each dataset.
```{r}
# Unique values in the country_txt column of terrorism_data
unique_terrorism_countries <- unique(terrorism_data$country_txt)
print("Unique values in terrorism_data$country_txt:")
print(unique_terrorism_countries)

# Unique values in the country column of poverty_data
unique_poverty_countries <- unique(poverty_data$country)
print("Unique values in poverty_data$country:")
print(unique_poverty_countries)
```

## Compare Countries
Create and execute an SQL query to compile a table that shows which countries are in each dataset. This helps to visually compare the datasets and see the overlap.
```{r}
# Create a table showing which dataset each country appears in
comparison_query <- "
SELECT country_txt AS country, 'Terrorism' AS source
FROM terrorism_data
UNION
SELECT country AS country, 'Poverty' AS source
FROM poverty_data
"

# Execute the query
comparison_table <- sqldf(comparison_query)

# View the comparison table
print(comparison_table)
```

## Finding Common Countries
Executes an SQL query to find countries that are present in both the terrorism and poverty datasets. The result provides a list of these common countries.
```{r}
# Query to find common countries in both datasets
query_common_countries <- "
SELECT DISTINCT country_txt
FROM terrorism_data
WHERE country_txt IN (SELECT DISTINCT country FROM poverty_data)
"

# Execute the query
common_countries <- sqldf(query_common_countries)

# Print results
print(common_countries)
```

## Creating Common Country DataFrame
This query retrieves records where both datasets have data for the same country and year. It includes columns from both datasets, allowing you to analyze the relationship between terrorism incidents and poverty metrics over time.
```{r}
# Query to find common countries and years in both datasets
# Query to find common countries and years in both datasets with additional variables
query_common_country_year <- "
SELECT DISTINCT
  terrorism_data.country_txt AS country,
  terrorism_data.iyear AS year,
  terrorism_data.attacktype1_txt,
  terrorism_data.targtype1_txt,
  terrorism_data.targsubtype1_txt,
  poverty_data.headcount_ratio_international_povline,
  poverty_data.headcount_ratio_lower_mid_income_povline,
  poverty_data.headcount_ratio_upper_mid_income_povline,
  poverty_data.headcount_international_povline,
  poverty_data.headcount_lower_mid_income_povline,
  poverty_data.headcount_upper_mid_income_povline
FROM
  terrorism_data
INNER JOIN
  poverty_data
ON
  terrorism_data.country_txt = poverty_data.country AND
  terrorism_data.iyear = poverty_data.year
"

# Execute the query
common_country_year <- sqldf(query_common_country_year)

# View the common countries, years, and additional variables
print(head(common_country_year))
```

# Data Analysis
## Attack Type Rank
This query counts the number of attacks for each type of attack and pivots the data for easier analysis. The result is a data frame where each row represents a country and columns show the count of different attack types.
```{r}
query_attacktype_count <- "
SELECT
  country AS country_name,
  attacktype1_txt AS attack_type,
  COUNT(*) AS attack_count
FROM
  common_country_year
GROUP BY
  country, attacktype1_txt
"
# Execute the query
attacktype_count_df <- sqldf(query_attacktype_count)

# Pivot the data
attacktype_pivot_df <- pivot_wider(attacktype_count_df, 
                                   names_from = attack_type, 
                                   values_from = attack_count, 
                                   values_fill = list(attack_count = 0))
print(head(attacktype_pivot_df))
```

Aggregates and plots the total count of each attack type across all countries. A horizontal bar plot is created to visually represent the frequency of different attack types.
```{r}
# Sum the attack counts for each attack type
attacktype_sums <- colSums(attacktype_pivot_df[,-1])  # Exclude the first column (country names)

# Convert to a data frame for plotting
attacktype_sums_df <- data.frame(
  attack_type = names(attacktype_sums),
  total_count = attacktype_sums
)

# Order by total_count for ranking
attacktype_sums_df <- attacktype_sums_df[order(attacktype_sums_df$total_count, decreasing = TRUE), ]

# Create the horizontal bar plot
ggplot(attacktype_sums_df, aes(x = reorder(attack_type, total_count), y = total_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for a horizontal bar plot
  labs(title = "Ranking of Attack Types by Frequency",
       x = "Attack Type",
       y = "Total Count of Attacks") +
  theme_minimal()
```

## Target Type Rank
Similar to the previous chunk, this query counts the number of attacks for each target type and pivots the data. This allows for analysis of different target types across countries.
```{r}
query_targettype_count <- "
SELECT
  country AS country_name,
  targtype1_txt AS target_type,
  COUNT(*) AS target_count
FROM
  common_country_year
GROUP BY
  country, targtype1_txt
"
# Execute the query
targettype_count_df <- sqldf(query_targettype_count)

# Pivot the data
targettype_pivot_df <- pivot_wider(targettype_count_df, 
                                   names_from = target_type, 
                                   values_from = target_count, 
                                   values_fill = list(target_count = 0))
print(head(targettype_pivot_df))
```

Calculates and plots the top 15 target types by frequency. The horizontal bar plot highlights which target types are most common.
```{r}
# Sum the attack counts for each attack type
targettype_sums <- colSums(targettype_pivot_df[,-1])  # Exclude the first column (country names)

# Convert to a data frame for plotting
targettype_sums_df <- data.frame(
  target_type = names(targettype_sums),
  total_count = targettype_sums
)

# Order by total_count for ranking
targettype_sums_df <- targettype_sums_df[order(targettype_sums_df$total_count, decreasing = TRUE), ]

# Create the horizontal bar plot with filtering in the plot code
ggplot(targettype_sums_df %>% 
         top_n(15, total_count) %>%  # Select the top 15 target sub types
         arrange(desc(total_count)),  # Ensure the top types are sorted in descending order
       aes(x = reorder(target_type, total_count), y = total_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for a horizontal bar plot
  labs(title = "Top 15 Target Types by Frequency",
       x = "Target Type",
       y = "Total Count of Target") +
  theme_minimal()
```

## Target Sub Type Rank
Counts attacks for each target subtype and pivots the data for further analysis. The result helps in understanding the distribution of different subtypes across countries.
```{r}
common_country_year$targsubtype1_txt[common_country_year$targsubtype1_txt == ""] <- "Unknown"

query_targetsubtype_count <- "
SELECT
  country AS country_name,
  targsubtype1_txt AS target_sub_type,
  COUNT(*) AS taget_sub_count
FROM
  common_country_year
GROUP BY
  country, targsubtype1_txt
"
# Execute the query
targetsubtype_count_df <- sqldf(query_targetsubtype_count)

# Pivot the data
targetsubtype_pivot_df <- pivot_wider(targetsubtype_count_df, 
                                   names_from = target_sub_type, 
                                   values_from = taget_sub_count, 
                                   values_fill = list(taget_sub_count = 0))
print(head(targetsubtype_pivot_df))
```

Calculates and plots the top 15 target subtypes by frequency. It includes data cleaning steps to handle character encoding issues and produces a horizontal bar plot similar to previous plots.
```{r}
# Sum the attack counts for each attack type
targetsubtype_sums <- colSums(targetsubtype_pivot_df[,-1])  # Exclude the first column (country names)

# Convert to a data frame for plotting
targetsubtype_sums_df <- data.frame(
  target_sub_type = names(targetsubtype_sums),
  total_count = targetsubtype_sums
)

# Order by total_count for ranking
targetsubtype_sums_df <- targetsubtype_sums_df[order(targetsubtype_sums_df$total_count, decreasing = TRUE), ]

# Convert the target_sub_type to UTF-8 encoding
targetsubtype_sums_df$target_sub_type <- iconv(targetsubtype_sums_df$target_sub_type, from = "latin1", to = "UTF-8")

# Create the horizontal bar plot with filtering in the plot code
ggplot(targetsubtype_sums_df %>% 
         top_n(15, total_count) %>%  # Select the top 15 target sub types
         arrange(desc(total_count)),  # Ensure the top types are sorted in descending order
       aes(x = reorder(target_sub_type, total_count), y = total_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for a horizontal bar plot
  labs(title = "Top 15 Target Sub Types by Frequency",
       x = "Target Sub Type",
       y = "Total Count of Target") +
  theme_minimal()
```

## Calculate Average poverty Ratios
Calculates the average poverty ratios and number of attacks per country and merges this data with previously prepared attack type, target type, and target subtype data frames to create a comprehensive dataset (final_df).
```{r}
# Query to calculate averages
query_country_povline_avg <- "
SELECT
  country AS country_name,
  COUNT(*) AS number_of_attack,
  AVG(headcount_ratio_international_povline) AS avg_international_povline_ratio,
  AVG(headcount_international_povline) AS avg_international_povline
FROM
  common_country_year
GROUP BY
  country
"

# Execute the query
country_povline_avg <- sqldf(query_country_povline_avg)

# List of data frames to merge
dfs <- list(country_povline_avg, attacktype_pivot_df, targettype_pivot_df, targetsubtype_pivot_df)

# Merge all data frames in the list by "country_name"
final_df <- Reduce(function(x, y) merge(x, y, by = "country_name"), dfs)
```

## Plotting Poverty Line Ratio with Num of Terrorist Attack
This scatter plot shows the relationship between the average international poverty line ratio and the number of terrorist attacks per country. A line is added to highlight trends.
```{r}
# Scatter plot of poverty ratio vs. number of terrorist attacks with a line
ggplot(final_df, aes(x = avg_international_povline_ratio, y = number_of_attack)) +
  geom_point() +
  geom_line() +
  labs(title = "Poverty Line vs. Terrorism Attacks",
       x = "International Poverty Line Ratio",
       y = "Number of Terrorism Attacks")
```

## Plot Terrorism Attack per Country in Real Map
Loads geographical data for countries and merges it with final_df to prepare for visualization on a map. The merge aligns country names between datasets.
```{r}
# Load the world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Ensure the country names are consistent
final_df$country_name <- as.character(final_df$country_name)

# Merge the data
world_merged <- left_join(world, final_df, by = c("name" = "country_name"))
```

Creates a map showing the number of terrorist attacks by country using the merged data. The map uses color gradients to represent attack counts.
```{r}
# Plot the map
ggplot(data = world_merged) +
  geom_sf(aes(fill = number_of_attack)) +
  scale_fill_viridis_c(option = "magma", direction = -1, na.value = "white") +
  labs(title = "Number of Terrorist Attacks by Country",
       fill = "Number of Attacks") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

# Save Dataset
This final chunk saves the merged dataset to a new CSV file for future use or sharing. This file contains all the combined insights from the terrorism and poverty datasets.
```{r}
# Save the new dataset to a CSV file
write.csv(final_df_fixed, "D:/TerrorismxPoverty.csv", row.names = FALSE, fileEncoding = "UTF-8")
```

```{r}
write.csv(country_povline_avg, "D:/country_povline_avg.csv", row.names = FALSE, fileEncoding = "UTF-8")
write.csv(attacktype_pivot_df, "D:/attacktype_pivot_df.csv", row.names = FALSE, fileEncoding = "UTF-8")
write.csv(targettype_pivot_df, "D:/targettype_pivot_df.csv", row.names = FALSE, fileEncoding = "UTF-8")
write.csv(targetsubtype_pivot_df_fixed, "D:/targetsubtype_pivot_df.csv", row.names = FALSE, fileEncoding = "UTF-8")

#(country_povline_avg, attacktype_pivot_df, targettype_pivot_df, targetsubtype_pivot_df)
```

```{r}
# Function to clean encoding issues in text
clean_encoding <- function(text) {
  # Replace problematic characters
  text <- iconv(text, from = "UTF-8", to = "ASCII//TRANSLIT", sub = "?")
  return(text)
}

# Apply the function to column names and data
colnames(targetsubtype_pivot_df) <- sapply(colnames(targetsubtype_pivot_df), clean_encoding)

# Check and fix any problematic entries in the data frame
targetsubtype_pivot_df[] <- lapply(targetsubtype_pivot_df, function(x) {
  if (is.character(x)) {
    clean_encoding(x)
  } else {
    x
  }
})
```

```{r}
targetsubtype_pivot_df_fixed <- data.frame(lapply(final_df, function(x) {
  if (is.character(x)) {
    clean_encoding(x)
  } else {
    x
  }
}), stringsAsFactors = FALSE)

colnames(targetsubtype_pivot_df_fixed) <- sapply(colnames(targetsubtype_pivot_df), clean_encoding)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
